---
title: '"mCP Packager Vignette"'
author: '"Hugo Amedei"'
date: "2023-05-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview 

The `mCP` package provides a set of functions for targeted protein complex detection in mass spectrometry-based proteomics experiments. It is designed to work with experimental data in the form of protein abundance matrices, and to compare the data to a database of known protein complexes to identify complexes that are present in the experimental data.

This vignette provides a step-by-step guide to using the `mCP` package to analyze proteomics data and identify protein complexes of interest.

# Installation

To install the mCP package from GitHub, run the following code:
```{r pressure, echo=FALSE}
library(devtools)
devtools::install_github("hugoagno3/mCP")
```


# Data input (two files Corum database and Experimental wide format results)

To perform targeted protein complexes detection, we need two input files:
           1- A CORUM database of protein complexes (or a targeted list of interest) dataframe (3 columns names:  col1= "complex_id", col2= "complex_name" and col3= "protein_id").
           2- An experiment file a *data.frame* with your experiment results in wide-format- first column called "protein_id" and the next columns can be numerics names from 1 to the last number of fractions, for example 1,2,3,..,35 for 35 fractions, means 35 columns. In this example, we will use the Corum_Humans_Database.csv file as the protein complex database and the Hek293_P2_1.csv and Hek293_P2_2.csv files as the experiment files.

```{r pressure, echo=FALSE}
library(readr)

# Read the Corum protein complex database file
Corum_Humans_Database <- read_csv("~/Documents/m_CP_Analysis/Documents to do DATA Analysis with m_CP/Hummans with new functions/Corum_Humans_Database.csv")

# Read the experiment files
Hek293_P2_1 <- read_csv("~/Documents/m_CP_Analysis/Mcp_paper/Diann_results/dDIA_DIANN/report.pg_matrix_HEK293_P2.tsv")
Hek293_P2_2 <- read_csv("~/Documents/m_CP_Analysis/Mcp_paper/Diann_results/dDIA_DIANN/report.pg_matrix_HEK293_P2.tsv")

# Select the matrix with columns representing protein intensities in each fraction from FR1 to FR35
Hek293_P2_1 <- Hek293_P2_1 %>% select(1, 6:75)
Hek293_P2_2 <- Hek293_P2_2 %>% select(1, 76:145)

```

# Data Processing (Input of matrix with long names wide format)

To process the input data, we need to run the mCP function or 3 functions provided by our mCP package.
# Option 1: Running mCP function
   This function is an integrated function of mCP package, that needs as input an experimental data and returns a list of plots, binary total hits, id of proteins of binary hits and heatmaps_seaborn of know protein complexes detected in CORUM database. In addition, it plots 4 files as  outputs: 
 1- pdf file with detected protein complexes profiles from Corum database.
 2- pdf with heatmaps of the detected protein complexes.
 3- txt file with numbers about general false positive when just 1 hit is consider as filter.
 4- CVS file containing all protein complexes detected, hits of binary interactions inside the protein complexes, FDR detected by MonteCarloSimulation.
   An example can be found here:
 
```{r pressure, echo=FALSE}
library(mCP)

##### Example #####
out_Hek_P2_1_teste <- mCP(corum_database = Corum_Humans_Database,
                          experiment_data = Hek293_P2_1, 
                          N_fractions = 35, 
                          specie = "hsapiens",
                          method_cor = "pearson",
                          heatmap_seaborn = TRUE,
                          format = "pdf", 
                          output_name = "m_CP_analysis_2",
                          filter = 0.93,
                          heat_map = TRUE,
                          relative = FALSE,
                          n_simulations= 9,
                          display_weights = TRUE,
                          standard_weights = list(list(x =11, label= "1049KDa"), 
                                             list(x = 13, label ="720 KDa")))
```
 
# Option 2: Runn mCP function individually 
```{r pressure, echo=FALSE}
library(mCP)

# Calculate the mean intensity of the replicates
NAmatrix_P2_1 <- read.table("/home/hugo/Documents/m_CP_Analysis/Mcp_paper/Diann_results/dDIA_DIANN/P2/Hek293_P2_1.csv",sep =",",dec = ".", header= T)
Hek_293_MA_P2_1 <- Calc_mean_matrix(NAmatrix = NAmatrix_P2_1,
                                     pattern_group_A = "_A_",
                                     pattern_group_B = "_B_",
                                     frac_index = 17,
                                     Protein_ID_column = 1,
                                     save_file = TRUE,
                                     save_name = "Hek293_P2_1.csv")
```

# Data Processing (Input of matrix with protein_id as first column and Factions)
  You can also go into the mCP procesion function with a wide format that 1st column is named "protein_id" and the other columns cointain the intensities of 

```{r pressure, echo=FALSE}

# Create a protein complex list based on the Corum database and the experimental results
CL_hek_P2_1 <- mcp_list(corum_database = Corum_Humans_Database,
                        experiment_data = Hek_293_MA_P2_1,
                        N_fractions = 35,
                        specie = "hsapiens")

###these are examples of the plots RUN
CL_hek_P2_1$`anti-BHC110 complex`

```

# First detection of Protein complex detection
  cpp_ploter filter protien complexes that at least have a binary inteaction higher that the filter value (Pearson correlation) and that are present in CORUM data base

```{r pressure, echo=FALSE}

out_Hek_P2_1 <- cpp_plotter(complex_list = CL_hek_P2_1, format = "pdf", output_name = "HEK293_P2_1_DIANN_dDIA", filter = 0.93, N_fractions = 35,
                            heat_map = TRUE, relative = FALSE,   display_weights = TRUE, standard_weights = list(list(x =11, label= "1049KDa"), 
                   list(x = 13, label ="720 KDa"),
                   list(x = 17, label = "480 KDa"), 
                   list(x =22, label ="146 KDa"),
                   list(x =29, label ="60 KDa"))
```

# Data Input and Pre-processing


# Summary of Package Functions

The `mCP` package provides the following functions:

- `Calc_mean_matrix()`: pre-processes the data before running `mcp_list()`.

- `mcp_list()`: creates a list of protein complexes based on CORUM database (complexes and experimental data).

- `cpp_plotter()`: filters the list of protein complexes based on the presence of co-eluting proteins calculates as a minimun of 1 binary interactions higher that the filter of protien within the candidate protein complex and creates complexome profiling plots, heatmaps and networks plots of proteins with in protein complexes of the selected complexes.

####  dd

Before using the `mCP` package, it is necessary to prepare the experimental data for analysis. The input data should be a protein abundance matrix in which the first row in called "protein_id" and the the others columns represents the intensity detected for each fraction of the protein complexes. The first column should contain the protein_id, and the remaining columns should contain the protein abundance values. these are numeric names from Fr 1 to the last one of the co-fractionation experiment

mCP R package is focused on detection of protein complexes and there fore it acepts only protein data as an imput matrix. The mass spectrometry data aquisition can be done by Data Dependent Acquisition mode or Data independent Aquisition.  The following steps are recommended for pre-processing the data before getting into mCP:

1. Normalize the data between fractions in experiments.  
2. Replace missing values by 0.
2. For comparison the normalization of the data to plot complexome profiling plots can be done inside each protein complexe by activateion the relativization fuction relative= TRUE in ccp_plotter function. 
3. Normalize per Biologica replicates if fractionation is reproducible. 


# Using the mcp_list() Function

The `mcp_list()` function takes two arguments: a database of known protein complexes and the experimental data in the form of a protein abundance matrix. The function returns a list of protein complexes that are present in the experimental data.

Here is an example of how to use the `mcp_list()` function:









Data

To perform targeted protein complexes detection, we need two input files: a protein complex database file and an experiment file. In this example, we will use the Corum_Humans_Database.csv file as the protein complex database and the Hek293_P2_1.csv and Hek293_P2_2.csv files as the experiment files.

{r

library(readr)

# Read the Corum protein complex database file
Corum_Humans_Database <- read_csv("~/Documents/m_CP_Analysis/Documents to do DATA Analysis with m_CP/Hummans with new functions/Corum_Humans_Database.csv")

# Read the experiment files
Hek293_P2_1 <- read_csv("~/Documents/m_CP_Analysis/Mcp_paper/Diann_results/dDIA_DIANN/report.pg_matrix_HEK293_P2.tsv")
Hek293_P2_2 <- read_csv("~/Documents/m_CP_Analysis/Mcp_paper/Diann_results/dDIA_DIANN/report.pg_matrix_HEK293_P2.tsv")

# Select the matrix with columns representing protein intensities in each fraction from FR1 to FR35
Hek293_P2_1 <- Hek293_P2_1 %>% select(1, 6:75)
Hek293_P2_2 <- Hek293_P2_2 %>% select(1, 76:145)

Data Processing

To process the input data, we need to run several functions provided by the mCP package.

{r

library(mCP)

# Calculate the mean intensity of the replicates
NAmatrix_P2_1 <- read.table("/home/hugo/Documents/m_CP_Analysis/Mcp_paper/Diann_results/dDIA_DIANN/P2/Hek293_P2_1.csv",sep =",",dec = ".", header= T)
Hek_293_MA_P2_1 <- Calc_mean_matrix(NAmatrix = NAmatrix_P2_1,
                                     pattern_group_A = "_A_",
                                     pattern_group_B = "_B_",
                                     frac_index = 17,
                                     Protein_ID_column = 1,
                                     save_file = TRUE,
                                     save_name = "Hek293_P2_1.csv")

# Create a protein complex list based on the Corum database and the experimental results
CL_hek_P2_1 <- mcp_list(corum_database = Corum_Humans_Database,
                        experiment_data = Hek_293_MA_P2_1,
                        N_fractions = 35,


```r
# Load the mCP package
library(mCP)

# Load the Corum database
Corum_Humans_Database <- read_csv("~/Documents/m_CP_Analysis/Documents to do DATA Analysis with m_CP/Hummans with new functions/Corum_Humans_Database.csv")

# Load the experimental data
NAmatrix_P2_1 <- read.table("/home/hugo/Documents/m_CP_Analysis/Mcp_paper/Diann_results/dDIA_DIANN/P2/Hek293_P2_1.csv", sep = ",", dec = ".", header = T)

# Pre-process the data
Hek_293_MA_P2_1 <- Calc_mean_matrix(NAmatrix = NAmatrix_P






